---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->
<!-- used devtools::build_readme() to update the md -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "",
  dev = "svg"
)
```

<!-- badges: start -->

[![CRAN status](https://www.r-pkg.org/badges/version/rnaturalearth)](https://CRAN.R-project.org/package=rnaturalearth)
[![](https://badges.ropensci.org/22_status.svg)](https://github.com/ropensci/software-review/issues/22)
[![Project Status: Active – The project has reached a stable, usable state and is being actively developed.](https://www.repostatus.org/badges/latest/active.svg)](https://www.repostatus.org/#active)
[![R-CMD-check](https://github.com/ropensci/rnaturalearth/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/ropensci/rnaturalearth/actions/workflows/R-CMD-check.yaml)

<!-- badges: end -->

# rnaturalearth

An R package to hold and facilitate interaction with [Natural Earth](https://www.naturalearthdata.com/) map data.

This package provides :

- access to a pre-downloaded subset of Natural Earth v4.1.0 (March 2018) vector data commonly used in world mapping

- easy subsetting by countries and regions

- functions to download other Natural Earth vector and raster data

- a simple, reproducible and sustainable workflow from Natural Earth data to rnaturalearth enabling updating as new versions become available

- clarification of differences in world maps classified by countries, sovereign states and map units

- consistency with Natural Earth naming conventions so that rnaturalearth users can use Natural Earth documentation

- data in 'sf' or 'sp' formats

The [Natural Earth](https://www.naturalearthdata.com/) website structures vector data by scale, category and type. These determine the filenames of downloads. rnaturalearth uses this structure to facilitate download (like an API).

## Install rnaturalearth

Install from CRAN :

```r
install.packages("rnaturalearth")
```

or install the development version from GitHub using [devtools](https://github.com/r-lib/devtools).

```r
devtools::install_github("ropensci/rnaturalearth")

```

Data to support much of the package functionality are stored in two data packages that you will be prompted to install when required if you do not do so here.

```r
devtools::install_github("ropensci/rnaturalearthdata")
devtools::install_github("ropensci/rnaturalearthhires")
```

## First usage

Here using `sp::plot` as a simple, quick way to plot maps. Maps could also be made with `ggplot2`, `tmap` or other options. All retrieval functions accept an argument `returnclass = "sf"` to return package `sf` (Simple Features) objects.

```{r, eval=TRUE}
library(rnaturalearth)
library(sp)

# world countries
sp::plot(ne_countries())

# uk
sp::plot(ne_countries(country = "united kingdom"))

# states, admin level1 boundaries
sp::plot(ne_states(country = "spain"))
```

## Introductory vignette

```{r, eval=FALSE}
vignette("rnaturalearth", package = "rnaturalearth")
```

## To download Natural Earth data not already in the package

There are a wealth of other data available at the [Natural Earth](https://www.naturalearthdata.com/) website. `rnaturalearth` has functions to help with download of these data.

The data available are outlined in the two tables below and online [here](https://www.naturalearthdata.com/downloads/50m-physical-vectors/).

```{r, eval=TRUE, echo=FALSE}

for (category in c("cultural", "physical")) {
  df_data <- read.csv(
    system.file(
      "extdata",
      paste0("data_list_", category, ".csv"),
      package = "rnaturalearth"
    )
  )
  cat("\ncategory  ", category, "\n")

  df_data
}
```

Specify the `scale`, `category` and `type` of the vector you want as in the examples below.

```{r, eval=FALSE}
# lakes
lakes110 <- ne_download(scale = 110, type = "lakes", category = "physical")
sp::plot(lakes110)

# rivers
rivers50 <- ne_download(
  scale = 50,
  type = "rivers_lake_centerlines",
  category = "physical",
  returnclass = "sf"
)

library(ggplot2)
library(sf)

ggplot(rivers50) +
  geom_sf() +
  theme_minimal()
```

## Details of different country definitions and scales

```{r, eval=FALSE}
vignette("what-is-a-country", package = "rnaturalearth")
```

## Reproducible download of Natural Earth data into the package

[Script](https://github.com/ropensci/rnaturalearthdata/blob/master/data-raw/data_download_script.r) used to get data into the accompanying data packages.

## Acknowledgements

Thanks to [Lincoln Mullen](https://github.com/lmullen) for code structure inspiration from [USAboundaries](https://github.com/ropensci/USAboundaries), [Hadley Wickham](https://github.com/hadley) for comments and prompting, [Bob Rudis](https://github.com/hrbrmstr) for answers to stackoverflow questions about downloading Natural Earth data into R. The [Natural Earth team](https://www.naturalearthdata.com/about/contributors/) and [Nathan Kelso](https://github.com/nvkelso) for providing such a great resource.

## Potential future work

### Potential additional data

- Country synonyms lookup
  - dataframe with ISO3 and country synonyms
  - similar to https://github.com/AndySouth/rworldmap/blob/master/data/countrySynonyms.rda
- Country larger regions lookup
  - dataframe with ISO3 and membership of different regional groupings, e.g. continent, least developed countries etc.
  - similar to https://github.com/AndySouth/rworldmap/blob/master/data/countryRegions.rda

### Potential additional functions

- facilitate joining of user data to country boundaries

  - similar to https://github.com/AndySouth/rworldmap/blob/master/R/joinCountryData2Map.R
  - ... but with a better name
  - similar allowing of join by ISO codes or names, with attempted synonym matching
  - similar reporting of country joining success and failure

- facilitate subsetting by country groupings
  - e.g. least developed countries etc.


