---
title: "Finding data"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Finding data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(rnaturalearth)
library(sf)
```

## Available data

There are a lot of data that can be downloaded from [Natural Earth](https://www.naturalearthdata.com/) with `ne_download()`. These data are divided into two main categories: *physical* and *cultural* vector data. The `df_layers_physical` and `df_layers_cultural` data frames included in the `rnaturalearth` packages show what layer of data can be downloaded.

### Physical vector data

```{r, eval=TRUE}
data(df_layers_physical)

knitr::kable(
  df_layers_physical,
  caption = "physical vector data available via ne_download()"
)
```

Based on the previous table, we know that we can download the `ocean` vector at small scale (110). Note that scales are defined as one of `110`, `50`, `10` or `small`, `medium`, `large`.

```{r}
plot(
  ne_download(type = "ocean", category = "physical", scale = "small")["geometry"],
  col = "lightblue"
)
```

### Cultural vector data

```{r, eval=TRUE}
data(df_layers_cultural)

knitr::kable(
  df_layers_cultural,
  caption = "cultural vector data available via ne_download()"
)
```

```{r}
plot(
  ne_download(
    type = "airports",
    category = "cultural",
    scale = 10
  )["geometry"],
  pch = 21,
  bg = "grey"
)
```

## Searching for countries and continents

In this article, we explore how we can search for data available to download within `rnaturalearth`. Let's begin by loading country data using the `read_sf()` function from the `sf` package. In the following code snippet, we read the Natural Earth dataset, which contains information about the sovereignty of countries.


```{r}
df <- read_sf("/vsizip/vsicurl/https://www.naturalearthdata.com/http//www.naturalearthdata.com/download/10m/cultural/ne_10m_admin_0_sovereignty.zip")

head(df)
```

### Finding countries

One way to search for countries is to search within the `ADMIN` vector. Let's start by plotting some of the first countries.

```{r, echo=-1, fig.show="hold", out.width="50%", results='hide'}
par(mfrow = c(2, 3))
lapply(df$ADMIN[1:6], \(x) {
  plot(ne_countries(country = x)["geometry"], main = x)
})
```

Suppose that we want to search the polygons for the US, how should we spell it?

```{r, eval=FALSE}
ne_countries(country = "USA")
ne_countries(country = "United States")
ne_countries(country = "United States Of America")
ne_countries(country = "United States of America")
```

One possibility consists to search within the `ADMIN` vector using a regular expression to find all occurrences of the word *states*.

```{r}
df$ADMIN[grepl("states", df$ADMIN, ignore.case = TRUE)]
```
We can now get the data.

```{r}
plot(ne_countries(country = "United States of America")["geometry"])
```
### Continents

Finally, let's create plots for each continent using the `ne_countries` function with the continent parameter.

```{r}
unique(df$CONTINENT)
```

```{r, echo=-1, fig.show="hold", out.width="50%", results='hide'}
par(mfrow = c(2, 4))
lapply(unique(df$CONTINENT), \(x) {
  plot(
    ne_countries(
      continent = x,
      scale = "medium"
    )["geometry"],
    main = x
  )
})
```
