---
title: "Language support in rnaturalearth"
author: "Andy South"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
#to produce a pdf
#output: rmarkdown::pdf_document
vignette: >
  %\VignetteIndexEntry{Language Support}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

(using languages but also about using with tmap and ggplot)

This vignette shows how [rnaturalearth](https://github.com/ropensci/rnaturalearth) makes it easier to make maps with labels in languages other than Engish, and to make thematic maps if you have data that is referenced in languages other than Engish.

[rnaturalearth](https://github.com/ropensci/rnaturalearth) is an R package to hold and facilitate interaction with natural earth vector map data.

[Natural Earth](https://www.naturalearthdata.com/) is a public domain map dataset including vector country boundaries. 


#### load required packages
```{r, eval=TRUE, echo=TRUE, message=FALSE}
library(rnaturalearth)
library(sp)
library(sf)
library(ggplot2)
library(ggrepel)
# library(tmap)
library(knitr)
```


### testing showing a table of available data
```{r echo = FALSE, results = 'asis'}

# category <- 'physical'

# df_data <- read.csv( system.file("extdata", paste0("data_list_", category, ".csv"), package = "rnaturalearth") )

# convert true,false to 0,1, but loses the name row
# I should just store as 0/1
# df_data <- as.data.frame(lapply(df_data[2:4],as.numeric))

knitr::kable(df_layers_physical, caption = "physical vector data available via ne_download()")

knitr::kable(df_layers_cultural, caption = "cultural vector data available via ne_download()")
```


### Country maps with labels in other languages

```{r, eval=FALSE, echo=TRUE, message=FALSE}
# eval FALSE while testing

# Africa
sp::plot(ne_countries(continent = "africa"))

sfaf <- ne_countries(continent = "africa", returnclass = "sf")
sfafc <- st_centroid(sfaf)

sfaf <- cbind(sfaf, st_coordinates(st_centroid(sfaf$geometry)))

# G = cbind(G, st_coordinates(st_centroid(G$geometry)))

# this adds centroids in the middle of countries
ggplot(sfaf) +
  geom_sf() +
  geom_sf(data = sfafc)

# trying labels in the middle of countries, doesn't quite work needs x,y,label
# but once x & y added on with st_coordinates ...
# seems getting the coords might not be necessary for much longer https://github.com/slowkow/ggrepel/issues/111
# cool nearly there ...
ggplot(sfaf) +
  geom_sf() +
  geom_text_repel(aes(x = X, y = Y, label = name))
# geom_text_repel(data=sfafc, aes(x=X, y=Y, label=name))
# geom_label_repel(data=sfafc, aes(x=X, y=Y, label=name))
# geom_text(data=sfafc, label='name_es', x='X', y='Y')

# getting there, labels still overlap a bit
# maybe make map bigger to allow space for labels
ggplot(sfaf) +
  geom_sf() +
  geom_text_repel(aes(x = X, y = Y, label = name_es))

# point.padding=NA allows labels to overlap the centroid
ggplot(sfaf) +
  geom_sf() +
  geom_text_repel(aes(x = X, y = Y, label = name_es), point.padding = NA)

# Africa labels just down left & right sides
# works pretty well I think
ggplot(sfaf) +
  geom_sf() +
  xlim(-28, 61) +
  geom_text_repel(aes(x = X, y = Y, label = name_es),
    data          = subset(sfaf, X > 21),
    nudge_x       = 60 - subset(sfaf, X > 21)$X,
    segment.size  = 0.2,
    segment.color = "grey50",
    direction     = "y",
    hjust         = 0
  ) +
  geom_text_repel(aes(x = X, y = Y, label = name_es),
    data          = subset(sfaf, X < 21),
    nudge_x       = -19 - subset(sfaf, X < 21)$X,
    segment.size  = 0.2,
    segment.color = "grey50",
    direction     = "y",
    hjust         = 1
  )

# french labels
ggplot(sfaf) +
  geom_sf() +
  xlim(-28, 61) +
  geom_text_repel(aes(x = X, y = Y, label = name_fr),
    data          = subset(sfaf, X > 21),
    nudge_x       = 60 - subset(sfaf, X > 21)$X,
    segment.size  = 0.2,
    segment.color = "grey50",
    direction     = "y",
    hjust         = 0
  ) +
  geom_text_repel(aes(x = X, y = Y, label = name_fr),
    data          = subset(sfaf, X < 21),
    nudge_x       = -19 - subset(sfaf, X < 21)$X,
    segment.size  = 0.2,
    segment.color = "grey50",
    direction     = "y",
    hjust         = 1
  )



# tmap good but labels currently overlap

# english labels
tm_shape(sfaf) +
  tm_borders() +
  tm_text("name")

# spanish labels
tm_shape(sfaf) +
  tm_borders() +
  tm_text("name_es")

# other languages de, fr, nl,
tm_shape(sfaf) +
  tm_borders() +
  tm_text("name_de", size = 0.5)

# chinese labels don't currently work, because attribute data is filled with NA
# I suspect I may need to do something different on reading the data in ?
# tm_shape(sfaf) +
#      tm_borders() +
#      tm_text("name_zh")

# The full list of
#    languages is: name_ar, name_bn, name_de, name_en, name_es, name_fr, name_el,
#    name_hi, name_hu, name_id, name_it, name_ja, name_ko, name_nl, name_pl,
#    name_pt, name_ru, name_sv, name_tr, name_vi, and name_zh.
# A 2-character language code decoder ring is here:https://en.wikipedia.org/wiki/List_of_ISO_639-2_codes.
```



