---
title: "rnaturalearth vignette"
author: "Andy South"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
#to produce a pdf
#output: rmarkdown::pdf_document
vignette: >
  %\VignetteIndexEntry{rnaturalearth vignette}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

This vignette is an introduction to [rnaturalearth](https://github.com/AndySouth/rnaturalearth), an R package to hold and facilitate interaction with natural earth vector map data.

[Natural Earth](http://www.naturalearthdata.com/) is a public domain map dataset including vector country and other administrative boundaries. 

[rnaturalearth](https://github.com/AndySouth/rnaturalearth) does two main things.

1. contains pre-downloaded vector maps for countries: `ne_countries()` and states: `ne_states()`
1. has `ne_download()` function to facilitate download of other vector maps


This vignette uses `sp::plot` as a simple, quick way to plot maps. More elaborate maps can be made with `rnaturalearth` data using `ggplot2`, `tmap` and other options.

#### load required packages
```{r, eval=TRUE, echo=TRUE, message=FALSE}
require(rnaturalearth)
require(sp)
```


## 1. maps in the package
Pre-downloaded maps can be accessed with `ne_countries()` for country (admin-0) boundaries and `ne_states()` for boundaries within countries (admin-1).

```{r, eval=TRUE, echo=TRUE, message=FALSE}

#world at small scale (low resolution)
sp::plot(ne_countries(type='countries', scale='small'))

#countries, UK undivided
sp::plot(ne_countries(country = 'united kingdom', type='countries'))
#map_units, UK divided into England, Scotland, Wales and Northern Ireland
sp::plot(ne_countries(country = 'united kingdom', type='map_units'))


#countries, large scale
sp::plot(ne_countries(country = 'united kingdom', scale='large'))

#countries, medium scale
#temporarily commented out to avoid utf conversion travis error with Curacao
#sp::plot(ne_countries(country = 'united kingdom', scale='medium'))
     
#countries, small scale
sp::plot(ne_countries(country = 'united kingdom', scale='small'))   

#states country='united kingdom'
sp::plot(ne_states(country ='united kingdom'))  
#states geounit='england'
sp::plot(ne_states(geounit='england')) 

#states country='france'
sp::plot(ne_states(country ='france'))  

```




## 2. downloading other Natural Earth vectors with ne_download()  

Each [Natural Earth](http://www.naturalearthdata.com/) dataset is characterised on the website according to `scale`, `type` and `category`. [rnaturalearth](https://github.com/AndySouth/rnaturalearth) allows you to specify `scale`, `type` and `category` and will construct the url and donload the corresponding file.

```{r, eval=TRUE, echo=TRUE, message=FALSE}

#lakes
lakes110 <- ne_download(scale=110, type='lakes', category='physical')
sp::plot(lakes110, col='blue')

#rivers
rivers110 <- ne_download(scale=110, type='rivers_lake_centerlines', category='physical')
sp::plot(rivers110, col='blue')


```


