---
title: "What is a country ?"
author: "Andy South"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
#to produce a pdf
#output: rmarkdown::pdf_document
vignette: >
  %\VignetteIndexEntry{What is a country ?}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

This vignette addresses potential confusion in mapping countries in [rnaturalearth](https://github.com/AndySouth/rnaturalearth).  
    
For example, from my own parochial perspective, if you are mapping the UK, do you wish to separate England, Scotland, Wales and Northern Ireland ? Do you wish to include the Falkland Islands. If you are mapping France, do you wish to include French Guiana ?

This vignette aims to help you.

[rnaturalearth](https://github.com/AndySouth/rnaturalearth) is an R package to hold and facilitate interaction with natural earth vector map data.

[Natural Earth](http://www.naturalearthdata.com/) is a public domain map dataset including vector country boundaries. 

This vignette uses `sp::plot` as a simple, quick way to plot maps. Similar maps could be made with `ggplot2`, `tmap` or other options.

#### load required packages
```{r, eval=TRUE, echo=TRUE, message=FALSE}
require(rnaturalearth)
require(sp)
```


### Country types : countries, map_units and sovereignty
```{r, eval=TRUE, echo=TRUE, message=FALSE}

#countries, UK undivided
sp::plot(world_countries(country = 'united kingdom', type='countries'))
#map_units, UK divided into England, Scotland, Wales and Northern Ireland
sp::plot(world_countries(country = 'united kingdom', type='map_units'))
#sovereignty, Falkland Islands included in UK
sp::plot(world_countries(country = 'united kingdom', type='sovereignty'), col='red')
sp::plot(coastline110, col='lightgrey', lty=3, add=TRUE)

#France, country includes French Guiana
sp::plot(world_countries(country = 'france'))
#France map_units includes French Guiana too
sp::plot(world_countries(country = 'france', type='map_units'))
#France filter map_units by geounit to exclude French Guiana
sp::plot(world_countries(geounit = 'france', type='map_units'))
#France sovereignty includes South Pacicic islands
sp::plot(world_countries(country = 'france', type='sovereignty'), col='red')
sp::plot(coastline110, col='lightgrey', lty=3, add=TRUE)



```

### Country scales : small, medium and large
```{r, eval=TRUE, echo=TRUE, message=FALSE}

#countries, large scale
sp::plot(world_countries(country = 'united kingdom', scale='large'))

#countries, medium scale
#temporarily commented out to avoid utf conversion travis error with Curacao
#sp::plot(world_countries(country = 'united kingdom', scale='medium'))
     
#countries, small scale
sp::plot(world_countries(country = 'united kingdom', scale='small'))     
     
     
```


### States, admin level 1, select by country or geounit 
```{r, eval=TRUE, echo=TRUE, message=FALSE}

#states country='united kingdom'
sp::plot(world_states(country ='united kingdom'))  
#states geounit='england'
sp::plot(world_states(geounit='england')) 

#states country='france'
sp::plot(world_states(country ='france'))  
#states geounit='france'
sp::plot(world_states(geounit='france')) 

```


