---
title: "What is a country ?"
author: "Andy South"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
#to produce a pdf
#output: rmarkdown::pdf_document
vignette: >
  %\VignetteIndexEntry{What is a country ?}
%\VignetteEngine{knitr::rmarkdown}
\usepackage[utf8]{inputenc}
---

This vignette addresses potential confusion in mapping countries in [rnaturalearth](https://github.com/AndySouth/rnaturalearth).  
    
For example, from my own parochial perspective, if you are mapping the UK, do you wish to separate England, Scotland, Wales and Northern Ireland ? Do you wish to include the Falkland Islands. If you are mapping France, do you wish to include French Guiana ?

This vignette aims to help you.

[rnaturalearth](https://github.com/AndySouth/rnaturalearth) is an R package to hold and facilitate interaction with natural earth vector map data.

[Natural Earth](http://www.naturalearthdata.com/) is a public domain map dataset including vector country boundaries. 



#### UK as an example
```{r, eval=TRUE, echo=TRUE, message=FALSE}

require(sp)
plot(world_countries(country = 'united kingdom', type='countries'))
plot(world_countries(country = 'united kingdom', type='map_units'))
plot(world_countries(country = 'united kingdom', type='sovereignty'))

```

#### France
```{r, eval=TRUE, echo=TRUE, message=FALSE}

require(sp)
plot(world_countries(country = 'france'))
#map_units includes French Guiana 
plot(world_countries(country = 'france', type='map_units'))
#aha need to filter by geounit to exclude (but doesn't work for type='countries')
plot(world_countries(geounit = 'france', type='map_units'))
plot(world_countries(country = 'france', type='sovereignty'))

```


